---
title: "Text as Data: Week 6"
author: "Matthias Haber"
date: "13 October 2021"
output:
    revealjs::revealjs_presentation:
      theme: moon
      reveal_plugins: ["notes", "zoom", "chalkboard"]
      highlight: haddock
      center: false
      self_contained: false
      incremental: true
      reveal_options:
        slideNumber: true
        progress: true
        previewLinks: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
library(tidyverse)
library(readtext)
library(quanteda)
```

# Goals for Today

## Goals

- Quanteda
- Text preprocessing
- Document-Feature Matrix

# Quanteda

## quanteda
The _quanteda_ package is an extensive text analysis suite for R, containing everything you need to perform a variety of automatic text analyses. _quanteda_ is to text analysis what dplyr and tidyr are to data wrangling.  

_quanteda_ is built for efficiency and speed, through its design around three infrastructures: the stringi package for text processing, the Matrix package for sparse matrix objects, and computationally intensive processing (e.g. for tokens) handled in parallelized C++.

In addition to the extensive documentation, there's a very helpful cheatsheet [here](https://muellerstefan.net/files/quanteda-cheatsheet.pdf)

```{r}
library(quanteda) #install.packages("quanteda")
```

## Build-in corpuses in quanteda
`quanteda` comes with several corpora included, like the corpus of US presidential inaugural addresses:

```{r}
corp <- quanteda::data_corpus_inaugural
summary(corp)
```


# Preprocessing

## Preprocessing Steps

Preprocessing text for analysis typically involves 8 steps:

1. Tokenizing the text to unigrams (or bigrams, trigrams, sentences, etc.)
2. Converting all characters to lowercase
3. Removing punctuation
4. Removing numbers & symbols
5. Removing stop words, including custom stop words
6. "Stemming" words, or lemmatization
7. Creating a Document-Feature Matrix
8. Removing sparse or frequent terms

## Tokenization 

The `token()` function from _quanteda_ allows to use external or the internal tokenizer to construct a token object, By default it creates word tokens and preserves hyphens, URLs, social media "tag" characters, and email addresses.

```{r}
txt <- c(text1 = "This is a short text,\n separated across two lines.",
         text2 = "Text-As-Data: https://github.com/mhaber/Text_as_Data_21")
tokens(txt)
```

## Tokenization

You can use the `what` argument to change the type of token object you want to create. For example, if you want it to construct sentences:

```{r}
tokens(c("Kurt Vonnegut said; only assholes use semi-colons.",
         "To be? Or not to be?"), what = "sentence")
```

## Tokenization

or characters:

```{r}
tokens("Just a bunch of characters.", what = "character")
```

## Tokenization

If you want to create n-grams in any length you can use the `tokens_ngrams()` function on a tokenized object.

```{r}
toks <- tokens(data_char_ukimmig2010[[3]])
tokens_ngrams(toks, n = 2:4) %>% head(10)
```

## Tokenization

You can also use other tokenizers like the ones contained in the `tokenizers` library.

```{r}
library(tokenizers)
song <-  paste0("How many roads must a man walk down\n",
                "Before you call him a man?\n",
                "How many seas must a white dove sail\n",
                "Before she sleeps in the sand?\n",
                "\n",
                "How many times must the cannonballs fly\n",
                "Before they're forever banned?\n",
                "The answer, my friend, is blowin' in the wind.\n",
                "The answer is blowin' in the wind.\n")
tokenizers::tokenize_paragraphs(song)
```

## Exercise

Tokenize the following tweet using quanteda's built-in word tokenizer, the tokenize_words tokenizer from the tokenizers package, and the tokenize_tweets tokenizer from the tokenizers package. What's the difference?

_https://t.co/8z2f3P3sUc  @datageneral FB needs to hurry up and add a laugh/cry button 😬😭😓🤢🙄😱 Since eating my feelings has not fixed the world's problems, I guess I'll try to sleep..._

## Conversion to lower case

In text analysis, it usually makes sense to convert characters to lower case although sometimes you may want to keep proper names capitalized. In `quanteda` you can use the `char_tolower()` directly on strings:

```{r}
test1 <- c(text1 = "England and France are members of NATO and UNESCO",
           text2 = "NASA sent a rocket into space.")
char_tolower(test1)
char_tolower(test1, keep_acronyms = TRUE)
```

## Conversion to lower case

or the `tokens_tolower()` function after you created your tokens object:

```{r}
test2 <- tokens(test1)
tokens_tolower(test2, keep_acronyms = TRUE)
```

## Removing punctuation, numbers & symbols

Removing unwanted features like punctuation, symbols and numbers is really easy and be done directly insight the `tokens` function:

```{r}
text <- "ph4t, phat, ph@	is 1337 for awesome or cool"
tokens(text, 
       remove_punct = TRUE, 
       remove_numbers = TRUE, 
       remove_symbols = TRUE)
```

## Removing stop words
As we've seen in the previous sessions, we often may want to remove stop words from our tokens object. `quanteda` automatically loads the `stopwords()` function from the `stopwords` package and defaults to the Snowball collection. Use `stopwords_getsources()` to get a list of available sources and `stopwords_getlanguages` to get a list of available languages for each source.

```{r}
stopwords::stopwords_getsources()
stopwords::stopwords_getlanguages("snowball")
```

## Removing stop words

We can remove stopwords from our tokens object with the `tokens_remove()` function. 

```{r}
text_token <- tokens("The quick brown fox jumps over the lazy dog")
tokens_remove(text_token, stopwords("en"))
```

## Removing other unwanted words

We can also use `tokens_remove()` to remove other unwanted words from our tokens object:

```{r}
tokens_remove(text_token, c("fox", "dog"))
```

## Exercise

Remove all the stopwords from the following Turkish poem:

"Yaşamak bir ağaç gibi
tek ve hür ve bir orman gibi   
kardeşçesine,
bu hasret bizim."

## Stemming

Stemming is the truncation of words to their root form, e.g. _playing_, _plays_, _played_ become _play_.

The tokenizers package provides a wrapper to the wordStem function from the SnowballC package, which applies a standard stemmer called the Porter stemmer. The stemmer is available for the following languages:

```{r}
SnowballC::getStemLanguages()
```

## Stemming

To reduce words to their word stem we can use the `tokens_wordstem()` function and apply it again to a tokens object:

```{r}
txt <- c(one = "eating eater eaters eats ate",
         two = "taxing taxes taxed my tax return")
th <- tokens(txt)
tokens_wordstem(th)
```

## All the preprocessing in one go

```{r}
# load data corpus
corp <- quanteda::data_corpus_inaugural
#create token
inaugural_tokens <- tokens(corp,
                       what = "word",
                       remove_punct = TRUE,
                       remove_symbols = TRUE,
                       remove_numbers = FALSE
                       )
# apply further preprocessing
inaugural_tokens_reduced <- inaugural_tokens %>% 
  tokens_tolower() %>%
  tokens_remove(stopwords('en')) %>% 
  tokens_wordstem()
```

## Keywords in context

A useful feature built into `quanteda` is keywords in context, which returns all the appearances of a word (or combination of words) in its immediate context.

```{r}
kwic(inaugural_tokens, "humble", window=3)
```

## Exercise

a) How many times has the word "slave" been used in inaugural addresses?
b) How many times has a word that included "slave" (like "slavery" or "enslaved") been used in inaugural addresses?

# Document-Feature-Matrix

## quanteda dfm

Quanteda is focused on bag-of-words (or bag-of-tokens) models that work from a document-feature matrix, where each row represents a document, each column represents a type (a "term" in the vocabulary) and the entries are the counts of tokens matching the term in the current document. 

To create a document-feature maxtrix use the `dfm()` function and apply it directly to the tokens object along with some common preprocessing options:

```{r}
my_dfm <- quanteda::dfm(inaugural_tokens,
                        tolower = TRUE)
```

## dfm summary

Typing the dfm's name will show an object summary.

```{r}
my_dfm
```

## dfm subsetting

You can look inside your dfm by indexing it like you would a Matrix object:

```{r}
my_dfm[1:5,1:5]
```

## dfm topfeatures

You can list the most (or least) frequently occuring features in a dfm by using the `topfeatures()` function:

```{r}
topfeatures(my_dfm, 40)
```

## dfm_trim()

You can also reduce the size of your dfm by removing sparse or frequently appearing terms with the `dfm_trim()` function. For example, to keep only those words that occur at least 10 times but not more than 100 times:

```{r}
dfm_trim(my_dfm, min_termfreq = 10, max_termfreq = 100)
```

## Exercise

Construct a `dfm` of trigrams from the inaugural speeches (lower case, no punctuation, not stemmed, no stop words removed).

a) How big is the matrix? How sparse is it?

b) What are the 20 most frequent trigrams?

c) Keep only those trigrams occurring at least 50 times and in at least 3/4 of the documents

# Wrapping up

## Questions?

## Outlook for our next session

- Next week there is no class!
- We'll meet again on October 27 where we'll learn different ways to visualize text
- I'll upload the code completion exercise this week. Please hand in your solution as an `.Rmd` file by October 26.

## That's it for today

Thanks for your attention!
